% arara: pdflatex: { synctex: yes }
% arara: makeindex: { style: ctuthesis }
%% arara: bibtex

%\listfiles


%\PassOptionsToPackage{cp1250}{inputenc}

% The class takes all the key=value arguments that \ctusetup does,
% and couple more: draft and oneside
\documentclass[twoside]{ctuthesis}

% MY IMPORTS
\usepackage{pdfpages}

\makeatletter
\edef\mytoday{\expandafter\@gobbletwo\the\year\ifnum\month<10 0\fi\the\month\ifnum\day<10 0\fi\the\day}
\makeatother

% LaTeX logo with better kerning in sf bf font
\makeatletter
\newcommand\LaTeX@lmss@bx{L\kern -.33em{\sbox \z@ T\vbox to\ht \z@ {\hbox {\check@mathfonts \fontsize \sf@size \z@ \math@fontsfalse \selectfont A}\vss }}\kern -.15em\TeX}
\DeclareRobustCommand\myLaTeX{%
	\ifcsname LaTeX@\f@family @\f@series\endcsname
		\csname LaTeX@\f@family @\f@series\endcsname
	\else
		\LaTeX
	\fi
}


\ctusetup{
	mainlanguage = english,
	secondlanguage = czech,
	otherlanguages = {czech},
%   title-czech = {Bias Detection in Czech News},
	title-english = {Bias Detection in Czech News},
	doctype = B,
	faculty = F3,
	department-czech = {Katedra počítačů},
	department-english = {Department of Computer Science},
	author = {Tomáš Horych},
	supervisor = {Ing. Jan Drchal, Ph.D},
	fieldofstudy-english = {Open Informatics},
	subfieldofstudy-english = {Artificial Intelligence and Computer Science},
	fieldofstudy-czech = {Matematcké inženýrství},
	subfieldofstudy-czech = {Matematické modelování},
	keywords-czech = {zpracování přirozeného jazyka, detekce zaujatosti, mediální bias, detekce subjektivity, klasifikace textu},
	keywords-english = {natural language processing, bias detection, media bias,subjectivity detection, text classification},
	day = 10,
	month = 2,
	year = 2022,
	specification-file = {my_modules/assignment.pdf},
	front-specification = true,
    pkg-listings = true,
    ctulstbg = none,
}

\ctuprocess

% Theorem declarations, this is the reasonable default, anybody can do what they wish.
% If you prefer theorems in italics rather than slanted, use \theoremstyle{plainit}
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[chapter]
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{conjecture}[theorem]{Conjecture}

\theoremstyle{note}
\newtheorem*{remark*}{Remark}
\newtheorem{remark}[theorem]{Remark}

% Marginpars used as navigation aids.
\usepackage{mparhack}
\usepackage{hyperref}
\usepackage[autostyle]{csquotes} 
\usepackage{subfig}
\usepackage{svg}
\usepackage{spverbatim}
\usepackage{fancyvrb}
\usepackage{fvextra}

% glossary input
\input{my_modules/gloss}
\usepackage{colortbl}
\definecolor{unbiased_clr}{RGB}{183,235,198}
\definecolor{biased_clr}{RGB}{236,189,189}
\definecolor{biasedw_clr}{RGB}{225,130,130}

\definecolor{highlight}{RGB}{198,220,240}


\newcommand\indexmp[1]{{\sffamily\bfseries#1}}

\ExplSyntaxOn
\cs_new:Nn \ctuman_domarginpar:n {
	\marginpar
	[ \raggedleft \footnotesize \sffamily #1 ]
	{ \raggedright \footnotesize \sffamily #1 }
}
\cs_generate_variant:Nn \ctuman_domarginpar:n { x }
\DeclareDocumentCommand \ctump { m } {
	\clist_set:Nn \ctuman_temp_clist { #1 }
	\ctuman_domarginpar:x { \clist_use:Nnnn \ctuman_temp_clist { \\ } { \\ } { \\ } }
	\clist_map_inline:Nn \ctuman_temp_clist { \index{##1|indexmp} }
	\ignorespaces
}
\ExplSyntaxOff



% Abstract in Czech
\begin{abstract-czech}
Automatická detekce zaujatosti v médiích představuje možnou cestu k objektivnějšímu a faktičtějšímu psaní. Tato práce se zaměřuje na zaujatost médií a zabývá se problémem binární klasifikace zaujatosti médií v českém zpravodajském prostředí. Je zde proveden přehled literatury a metodiky pro různé aspekty zaujatosti. Následně je shromážděn a analyzován soubor dat týkajících se zajuatosti médií. S využitím strojového překladu je prezentováno osm paralelních českých datasetů, přičemž jeden z nich je rozsáhlý dataset o 360 tisících větách. Kromě toho je z wikipedie automaticky extrahován nový český dataset CWNC (Czech Wiki Neutrality Corpus) pro detekci zaujatosti, s 5 766 větami. Následné experimenty pak ukazují vliv předtrénování na kombinacích aktuálních dostupných doménových datasetů a naznačují pozitivní vliv předtrénování na datasetech zaměřených na subjektivitu. Český BERT model s nejlepšími parametry a nastavením pak dosahuje skóre F1 \textbf{80,2\%} na vybraném testovacím dataset. Nakonec je natrénovaný klasifikátor využit ke klasifikaci reálných dat z různých zpravodajských zdrojů.

\end{abstract-czech}

% Abstract in English
\begin{abstract-english}
Automatic detection of bias in media represents a possible way to more objective and factual writing. This work focuses on \textbf{media bias} and addresses the problem of binary classification of media bias in the Czech news environment. The literature and methodology for different aspects of bias are reviewed. Then, a set of datasets related to media bias is collected and analyzed. Utilizing machine translation, eight parallel Czech datasets are presented, one of which being a large-scale dataset of 360k sentences. Additionally, a novel Czech dataset CWNC (Czech Wiki Neutrality Corpus) for bias detection with 5,766 sentences is automatically extracted from Wikipedia. Following experiments then show the effects of pre-training on combinations of current in-domain datasets, suggesting a positive effect of pre-training on subjective datasets. Czech BERT-based model with the best parameters and setting then achieves an F1 score of \textbf{80.2\%} on a selected target dataset. Finally, a trained classifier is utilized to classify real-world data from various news sources.

\end{abstract-english}

% Acknowledgements / Podekovani
\begin{thanks}
I thank to my supervisor, Jan Drchal, for the guidance through the Machine Learning and NLP methodology and for supporting my research interests. I would also like to thank to my family, who always supported me during my studies and throughout the process of work on this thesis.
\end{thanks}

% Declaration / Prohlaseni
\begin{declaration}
I declare that the presented work
was developed independently and that
I have listed all sources of information
used within it in accordance with the
methodical instructions for observing
the ethical principles in the preparation
of university theses.

\medskip

In Prague, \monthinlanguage{title} \ctufield{day}, \ctufield{year}

\medskip

\end{declaration}

\usepackage{url}
\usepackage{subfig}

\usepackage{tabularx,array}
\usepackage{mathtools,amssymb}
\usepackage{amsmath,amsthm,enumitem}
\usepackage{comment}


% A savebox for typesetting listings in the titles
\newsavebox{\myboxa}

%\newcommand*\symbO{$\color{red}\bowtie$}
\newcommand*\symbO{\raisebox{0.5\height}{\scalebox{0.7}{\color{red}${\vartriangleright}\mkern-6mu{\vartriangleleft}$}}}
\newcommand*\symbM{\raisebox{0.5\height}{\scalebox{0.7}{\color{red}${\blacktriangleright}\mkern-6mu{\blacktriangleleft}$}}}
\newcommand*\itemO{\item\leavevmode\kern-0.33em\symbO}
\newcommand*\itemM{\item\leavevmode\kern-0.33em\symbM}


\begin{document}

% We actually don't want inline listings to have a background color
\renewcommand \ctulstsep {0pt}

% \ctuclsname for typesetting the class' name
\newcommand\ctuclsname{\leavevmode\unhcopy\ctuclsnamebox}
\newsavebox\ctuclsnamebox
\begin{lrbox}{\ctuclsnamebox}
\ctulst!ctuthesis!
\end{lrbox}

% all the formal stuff (abstract, title page etc)

\maketitle















%-------------------------------------------------------------------------------------------------------------------
% MAIN TEXT
\input{my_modules/introduction}


\input{my_modules/sota}

\input{my_modules/datasets}

\input{my_modules/czech_datasets}

\input{my_modules/theoretical_background}

\input{my_modules/experiments}

\input{my_modules/inference}

\input{my_modules/Conclusion}





\printindex

\bibliographystyle{ieeetr}
\bibliography{my_modules/citation}
\printglossary
\appendix
\input{my_modules/AppendixA}
\input{my_modules/AppendixB}


\end{document}